\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-01-31 draft.tex
fix game-tree for uncertainty at trade war stage
move preference defns to new sec:pref

The timing is as follows. First, the executives set trade policy cooperatively within the context of an international agreement. After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.} Next, the uncertainty about the the median legislator's identity is resolved and the legislatures decide whether to abide by the agreement or to provoke a trade war. In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model. Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium. This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions.

\subsection{Preferences}
\label{sec:pref}

@@

For the case of trade policy, I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare. This alters none of the results and so would only serve to complicate the exposition.}}

@@

In this formulation, the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$. This allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks in the trade agreements literature. Alternatively, as motivated below, it can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature. Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.

>>>>

{\color{blue} The timing is as follows. First, the executives set trade policy cooperatively within the context of an international agreement. After the trade agreement is concluded, the lobbies attempt to persuade the legislators in their respective countries to vote ``no'' on the ratification of the agreement; or alternatively, if there is no ratification hurdle, to break the trade agreement.\footnote{As will be seen shortly, these two interpretations are equivalent in this setting because tariffs in a ``trade war'' are assumed to be set non-cooperatively in the same manner as if there were no trade agreement. For simplicity, I refer to this as the ``break'' decision throughout.} Next, the uncertainty about the the median legislator's identity is resolved and the legislatures decide whether to abide by the agreement or to provoke a trade war. In the event that the trade agreement does not enter into or remain in force, there is a final stage of lobbying, resolution of the uncertainty surrounding this decision, and voting to set the trade-war tariffs.\footnote{Alternatively, one can model the lobbying for the ratification decision as determining the potential non-cooperative tariff; this changes behavior slightly depending on the relationship between the lobby and the legislature's preferences, but does not change the fundamental dynamics of the model. Yet another possibility is to replace the trade-war phase with the ex-ante status quo, which would most naturally be the non-cooperative equilibrium. This is precisely the outcome that obtains in the trade-war phase, so the results of the model are unaltered.} Once all political decisions are taken, producers and consumers make their decisions. }

\subsection{Preferences}
\label{sec:pref}
I specify the welfare function of the home legislature as
\begin{equation}
  W_{\mathit{ML}} = \mathit{CS}_X(\tau) + \ga(e,\ve) \cdot \mathit{PS}_X(\tau) + \mathit{CS}_Y(\tau^*) + \mathit{PS}_Y(\tau^*) + \mathit{TR}(\tau)
  \label{eq:ml}
\end{equation}
where $\mathit{CS}$ is consumer surplus, $\mathit{PS}$ is producer surplus, $\ga(e,\ta)$ is the weight placed on producer surplus (profits) in the import-competing industry, $e$ is lobbying effort, $\ta$ is a random variable, and $\mathit{TR}$ is tariff revenue.\footnote{\textbf{Labor income $l$ could also be included in both median legislator and executive welfare. This alters none of the results and so would only serve to complicate the exposition.}} The weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$.

@@

For the case of trade policy, ...

@@

the weight the median legislator places on the profits of the import-competing industry, $\ga(e,\ve)$, is affected by the level of lobbying effort and a random variable $\ta$. This allows for the analysis of endogenous lobbying to be integrated with the standard political economy shocks in the trade agreements literature. Alternatively, as motivated below, it can represent reduced-form modeling of political uncertainty arising from a complex decision-making process within a legislature. Modeling the objective function so closely on the standard in the trade agreements literature allows for direct comparisons to the large extant body of work that studies exogenous shocks only, revealing cleanly the effects of the addition of endogenous lobbying---and if one wants---this different kind of political uncertainty.
